---
title: "0303_INCASI_preprocess_b"
author: "Stella"
date: "3 March 2018"
output: html_document
---


### INCASI setup

#### 1. General setup
```{r INCASI_setup, message = FALSE}
#--------------------------------------------------------------------------
# Clear workspace
rm(list=ls())

#--------------------------------------------------------------------------
# Set working directory to containing folder ('0303_INCASI'):

# knitr automatically sets this to the folder that contains the script 
# (here 0303_INCASI/_scripts)

#--------------------------------------------------------------------------
# Define paths:
path_wd 	 <- getwd()
path_data <- paste("../../0303_INCASI_data/",sep="") # path to folder: 0303_INCASI_data
path_scripts  <- paste("../_scripts/",sep="") # path to file: _scripts
path_functions  <- paste("../_functions/",sep="") # path to file: _functions
path_dataframes <- paste("../_dataframes/",sep="") # path to file: _dataframes
path_variables <- paste("../_variables/",sep="") # path to file: _variables
path_figures <- paste("../_figures/",sep="") # path to file: _figures

#--------------------------------------------------------------------------
# load packages
source(paste(path_scripts,"loadPackages.R",sep="")) # installed Rpackages

#--------------------------------------------------------------------------
# load settings for ggplot
source(paste(path_scripts,"loadTheme.R",sep="")) # ggplot theme
source(paste(path_scripts,"loadColors.R",sep="")) # ggplot colours

#--------------------------------------------------------------------------
# define list of participants
inc_list <- as.factor(dir(path_data)) #create inc_list with all inc04 - inc50 

#--------------------------------------------------------------------------
# specify ECG sampling frequency (in Hz)
ecg_sf <-  2048 # --> 1 datapoint every 0.4882812 ms (1000/2048)
```


####  2. Load dataframes for analysis: 
```{r scripts analysis, message = FALSE}
#----------------------------------------------------------------------------
# dataframes for analysis encoding 
load(paste(path_dataframes,"logfiles_all.RData", sep='')) # raw log_encode, log_recall, log_rate 
load(paste(path_dataframes,"main_exg2_peaks_s_list.RData", sep='')) # ecg lead with peaks (sec)

# load defined cardiac phases (created in templates_t_q)
load(paste(path_dataframes,"fin_sys_dur_s.RData", sep="")) # load systole lengths from Rpeak until twave end (created in templates_t_q)
load(paste(path_dataframes, "qonR_s.RData", sep="")) # load detected qwave-Rpeak lengths (created in templates_t_q)
#----------------------------------------------------------------------------
```


#### 3. Prepare supplementary analysis of picture ratings
* Load normative ratings from csv files
* `emopics_norms`: df with normative valence and arousal ratings for each picture (180 pictures)
```{r rating norm}
#----------------------------------------------------------------------------
# load normative data (norms_valence, norms_arousal) from EmoPics csv
rate_pos <- read.csv(file = paste(path_variables, "EmoPics_pos.csv", sep=""), header=TRUE, sep = ';', dec = ",")
rate_pos$valence <- "positiv"
rate_neu <- read.csv(file = paste(path_variables, "EmoPics_neu.csv", sep=""), header=TRUE, sep = ';', dec = ",")
rate_neu$valence <- "neutral"
rate_neg <- read.csv(file = paste(path_variables, "EmoPics_neg.csv", sep=""), header=TRUE, sep = ';', dec = ",")
rate_neg$valence <- "negativ"

emopics_norms <- rbind(rate_pos, rate_neu, rate_neg)

#save(emopics_norms, file = paste(path_dataframes,"emopics_norms.RData", sep="")) # save emopic ratings
#----------------------------------------------------------------------------
```


* Load experimental ratings from rating period
* `c_rate_stim`: df with individual mean valence and arousal ratings for each picture
* `rate_va`: joint dataframe of normative and individual ratings for arousal and valence
```{r rating individual}
#----------------------------------------------------------------------------
## experimental ratings: mean valence and arousal ratings for each pic 
m_rate_stim <- melt(log_rate, id = c("stimnum", "valence"), measure = c("rate_valence", "rate_arousal"))
c_rate_stim <- cast(m_rate_stim, stimnum * valence ~ variable, mean, na.rm = T)

# rate_va: join df with mean experimental + normative ratings for each picture
# valence: rate_valence, norms_valence
# arousal: rate_arousal, norms_arousal
rate_va <- left_join(c_rate_stim, emopics_norms[1:3], by = c('stimnum'))

# save(rate_va, file = paste(path_dataframes,"rate_va.RData", sep="")) # save rate_va
#----------------------------------------------------------------------------
```


#### 4. Prepare supplementary analysis of covariates
* Age
* Interoceptive Accuracy (IA)
* Trait anxiety (STAI)
* Heart Rate Variability (HRV: mean RR, rmssd)

```{r analyse covariates}
#--------------------------------------------------------------------------
# load dataframes
## age
data_age <- read.csv(file = paste(path_variables, "INCASI_age.csv", sep=""), header=TRUE, sep = ';', dec = ",")  # age df

## IA
load(paste(path_dataframes, "score.RData", sep="")) # hbp score df

## HRV
# a. read mean HR, rmssd, mean R-R (from main file = experimental period)
kubios_main <- read.csv(file = paste(path_variables, "data_kubios_main.csv", sep=""), header=TRUE, sep = ';', dec = ",")

# b. read mean HR, rmssd, mean R-R (from main file = resting period)
kubios <- read.csv(file = paste(path_variables, "data_kubios_rest.csv", sep=""), header=TRUE, sep = ';', dec = ",")

# save(kubios, file = paste(path_dataframes,"kubios.RData", sep="")) # save kubios

## STAI
stai <- data.frame(read.csv(file = paste(path_variables, "STAIT.csv", sep=""), header=TRUE, sep = ';', dec = ","))
stai$stai_sum <- rowSums(stai[2:length(stai)])
stai <- stai[!stai$vp == "inc48",] # exclude participant inc_48 (extrasystoles) to fit inc_list

# save(stai, file = paste(path_dataframes,"stai.RData", sep="")) # save stai

#--------------------------------------------------------------------------
# function flex_medsplit
# function to compute median split
# df: dataframe containing column to create median split on
# colname: name of column in df to compute median split on
# list: vector of participants
flex_medsplit <- function(df, colname, list, plot = F) {

  # 1. compute medsplit 
    vec <- as.numeric(df[[colname]]) # create vector (column named colname in df) to compute medsplit on
    median_vec <- median(vec, na.rm = T)  # calculate median of vector vec
    split = cut_number(vec, n = 2, na.rm = T, labels = c("below", "above")) # cut number of vector elements in equal groups, label them "above", "below"
    
  # 2. create dataframe medsplit  
    medsplit <- data.frame(
      vp = list,
      score = vec,
      medsplit = split
    )
    
    # 3. check if medsplit contains two equal parts
    if (length(split[split == "below"]) != length(split[split == "above"])) {
      message("medsplit is not equal:", "Below:", sum(vec < median_vec, na.rm=T), "Above:", sum(vec >= median_vec, na.rm=T))
      split[which(vec == median_vec)] <-  NA # exclude median_vec from medsplit factor
    }
  
  # 4. plot medsplit
    
    #assign each participant a number (for better readability x axis on plot)
    vplist <- unique(medsplit$vp)
    vpnum <- 0
    for(i in 1:length(vplist)){
    idx <- which(medsplit$vp==vplist[i])
     vpnum[idx] <- i
      }
   
  if (plot == T) {
    plot_medsplit <- ggplot(data = medsplit, aes(x= vpnum, y = vec, colour = medsplit)) +
      scale_colour_manual(values=c(defmedblue, deforange)) +
      geom_point(na.rm = T) +
      geom_abline(intercept = median_vec, slope = 0) +
      ggtitle(paste("Mediansplit", colname, "for: ", deparse(substitute(list))))+
      labs(x = "Participants", y = colname) +
      mytheme
    print(plot_medsplit)
  }
  
  # 5. return dataframe df_subs with column medsplit
  colnames(medsplit) <- c("vp", paste(colname, "_score", sep = ""), paste(colname, "_medsplit", sep = ""))
  return(medsplit)
}

#--------------------------------------------------------------------------
split_IA <- flex_medsplit(score, colname = "IA", inc_list, plot = T)
split_stai <- flex_medsplit(stai, colname = "stai_sum", inc_list, plot = T)
split_rmssd <- flex_medsplit(kubios, colname = "rmssd_rest", inc_list, plot = T)

data_var <- join_all(list(split_IA, split_stai, split_rmssd), by = c("vp")) # df containing all variables
#--------------------------------------------------------------------------
```


# 1. PREPARE LONG DATAFRAME: LOG_ENCODE

1.A. Determine systolic time intervals (STI) based on regression equations (Weissler, 1968)

1.B. Analysis of behaviour relative to heartbeat
**`log_encode`**: edit log_encode (one row per trial)
* `main_exg2_peaks_s_list`: create list with time points from R peaks for each participant
* `main_exg2_dtp`: create list with complete ecg leads for each participant


## 1.A. Compute Regression equations (Weissler, 1968) to determine Systolic time intervals

* Definition of STI (systolic time intervals)
* *QS2 = PEP + LVET*: the whole electromechanical systole, start: beginning of qRs complex (depolarization ventricles), end: closure of the aortic valve (end of systolic blood outflow)
* *PEP*: Pre-ejection period: itnerval from onset of ventricular depolarization to beginning of ventricular ejection
* *LVET*: phase of systole during which blood is ejected into the arterial sysytem (beginning of upstroke to trough of incisurila notch on external carotid arterial pulse)

* Systolic phases of cardiac cycle vary inversly with HR -> corrections for variation in heart rates required
* Regression equations relate each of the STI to the HR
* Most widely applied those by Weissler, Harris and Schoenfeld (1968):  recorded in normal subjects with wide range of resting HR

```{r function callsti}
# create function to calculate systolic time intervals (STI) based on Weissler regression equations (1968)

# function callsti
# x = mean participant HR in bpm
# df = dataframe (e.g. log_encode, log_recall)
# ivar = iteration variable (iclick, ipic)

callsti <- function (x, df, ivar) {
  gen <- df$gender[ivar] # gender
  if (gen == "m") { # STI for male
  lvet <- ((-1.7 * x) + 413) / 1000
  pep <- ((-0.4 * x) + 131) / 1000
  qs2 <- ((-2.1 * x) + 546) / 1000
  } else {  # STI for female
    lvet <- ((-1.6 * x) + 418) / 1000
    pep <- ((-0.4 * x) + 133) / 1000
    qs2 <- ((-2.0 * x) + 549) / 1000
    }
  sit_vec <- c(pep, lvet, qs2)
  names(sit_vec) <- c("pep", "lvet", "qs2")
  return(sit_vec)
}
```



## 1.B. Analysis of behaviour relative to heartbeat

### 1.B.1. Define relative timepoint within R-R interval, length of R-R interval, and heart rate for each key press

* What's the time of the Rpeak just before each key press (click1onset)?
```{r add var log_encode Rtend}
#--------------------------------------------------------------------------
#define variables for loop (if main_exg2_peaks_s, meain_exg2_dtp are not yet created) -> otherwise already loaded (see above)
# main_exg2_peaks_s <- NULL # create list to contain Rpeaks for each participant (vp)
# main_exg2_peaks_s_list <- NULL
main_exg2_dtp <- list() # create list to contain single ecg leads for each vp
print(getwd())
print(path_data)
#--------------------------------------------------------------------------
## 1. define relative timepoint, R-R interval, and heart rate for each click
for (part in inc_list[1:length(inc_list)]) {  # loop over participants
  
  cat('\014')
  print(paste('Processing subject', num2str(grep(part, inc_list), fmt=0), 'of',
               num2str(length(inc_list),fmt=0)))
  
  ## a) create main_exg_2 lists (if not created)
  # # Read single lead with dtp for each vp
  main_exg2_dtp[[part]] <- read.table(paste(path_data, part, '/', part, '_main_exg2.txt', sep = ''))
  # 
  # # Read time points from R peaks Kubios output
  # main_exg2_peaks_tmp <- NULL
  # main_exg2_peaks_tmp <- readMat(paste(path_data, part, '/', part, '_main_exg2_hrv.mat', sep = '')) # read Kubios matlab output into R
  # main_exg2_peaks_s <- as.vector(main_exg2_peaks_tmp$Res[[4]][[2]][[2]]) # extract tmp from Res structure in  matlab
  # main_exg2_peaks_s_list[[part]] <- main_exg2_peaks_s

  ## b) load main exg_2 (if already created)
  main_exg2_peaks_s <- main_exg2_peaks_s_list[[part]] # select list of Rpeaks for part
  
  #--------------------------------------------------------------------------
  run = min(which(log_encode$vp == part)) - 1 # create run variable to loop over respective rows for each vp
  for (iclick in (run+1):(run+length(log_encode$click1onset_ecgtime_s[log_encode$vp == part]))) {   # loop over 120 clicks for each vp

    #--------------------------------------------------------------------------
    # position of Rpeak before each key press (click1onset) timepoint
    clickon <- log_encode$click1onset_ecgtime_s[iclick] # timepoint of click
    log_encode$position[iclick] <-  max(which(main_exg2_peaks_s < clickon)) # select timepoints of Rpeaks right before clickonsets
    position <- log_encode$position[iclick] 
    
    #--------------------------------------------------------------------------
    # RR intervals of click (in sec): (position of Rpeak before click = position)
    log_encode$R_R_s[iclick] <- (main_exg2_peaks_s[position+1] - main_exg2_peaks_s[position]) # RR around click: calculate length of R-R-interval (one cardiac cycle)
    
    #--------------------------------------------------------------------------
    # compute beat to beat heart rate: R-R intervals, HR
    log_encode$HR_1perMin[iclick] <- (60 / log_encode$R_R_s[iclick]) # click to click heart rate: HR_1permin = 60 / diff.R-R (in bpm)

    log_encode$HR_1perSec[iclick] <- (1 / log_encode$R_R_s[iclick]) # heart rate: HR_1perSec = 1 / diff.R-R (in Hz)
    
    #--------------------------------------------------------------------------
    # for each clickonset: calculate difference to preceding R peak (in sec) -> create column diff_click_peak_s
    log_encode$diff_click_peak_s[iclick] <- (clickon - main_exg2_peaks_s[position]) 
  }
}    

#--------------------------------------------------------------------------
#save list with Rpeaks from Kubios output: 
#save(main_exg2_peaks_s_list, file = paste(path_dataframes,"main_exg2_peaks_s_list.RData", sep=''))

#save list with complete ecg leads for each vp:
#save(main_exg2_dtp, file = paste(path_dataframes,"main_exg2_dtp.RData", sep=""))
#--------------------------------------------------------------------------
```


### 1.B.2. Define circular onset and cardiac phase of each key press

Analyse key press relative to the heartbeat (1. circular and 2. binary analysis)

  1. Circular
  
  * Compute relative phase of click1onset within R-R interval, assigning values btw. 0 and 2pi (log_encode$radclick)
  
  2. Binary
  
  * In which cardiac phase do clicks fall - in systole (ejection-phase: sys1, sys2), diastole   (dias1, dias2, dias3, dias4), or non-defined (pre-ejection period, security window between systole and diastole)?


For binary analysis: Cardiac cycle is segmented into several intervals:
* qonR: interval from qwave onset until Rpeak (=qonR_s)
* crop: interval from Rpeak to end of pep (pep = qonR_s + crop)
* Rtend: Interval from Rpeak to twave end (here, end of systole)
* syspat: Rtend minus crop: ejection period until twave end (without pre-ejection) = lvet
* sys1: first half of ejection period (length: 0.5 * syspat)
* sys2: second half of ejection period (length: 0.5 * syspat)

* pep: pre-ejection period (starting from q wave)
* lvet: ventricular ejection time (qs2 - pep) = syspat

* diaspat: remaining diastole (R-R - syspat - qs2)
* dias1 to dias4: diastolic bins of same length (lenght: 0.25 * diaspat)

Two cardiac phases are defined as systole, diastole:
* 2 intervals in systole:  sys1, sys2 (ejection until twave end)
* 4 intervals in diastole: dias1, dias2, dias3, dias4
* non-defined cardiac intervals: pep (pre-ejection), 50ms window between systole and diastole

```{r add var log_encode Rtend}
#--------------------------------------------------------------------------    
## Run circular and binary analysis

for (part in inc_list[1:length(inc_list)]) {  # loop over participants

  run = min(which(log_encode$vp == part)) - 1 # create run variable to loop over respective rows for each vp
  for (iclick in (run+1):(run+length(log_encode$click1onset_ecgtime_s[log_encode$vp == part]))) {   # loop over 120 clicks for each vp
  
    #--------------------------------------------------------------------------
    dif <- log_encode$diff_click_peak_s[iclick] # vector dif with time interval from each click to preceding R (sec)

    # 1. CIRCULAR: transform dif (sec) into radian = relative phase of click1onset within cardiac cycle: 2 * pi * (HR_1permin/60) * (diff_click_peak_s)
    log_encode$radclick[iclick] <- (2 * pi * (log_encode$HR_1perSec[iclick]) * (dif)) 
    dif_rad <- log_encode$radclick[iclick] 

    #--------------------------------------------------------------------------
    # 2. BINARY: check in which cardiac intervals clicks fall
    
    # a) define systolic time intervals (STI)
    x <-  mean(log_encode$HR_1perMin[log_encode$vp == part])  # mean heart rate 
    sitvec <- callsti(x, log_encode, iclick) # vector with systolic intervals (calculated from qwave)
    
    log_encode$pep[iclick] <- sitvec["pep"] # pre-ejection period 
    log_encode$qonR[iclick] <- qonR_s[part] # difference from qonset to R
    
    # transfer STI to intervals that start from Rpeak (not qwave) -> substract distance qonR
    log_encode$crop[iclick] <- sitvec["pep"] - qonR_s[part] # crop: pre-ejection period from Rpeak, substract q
    log_encode$Rtend_s[iclick] <- sys_dur_s[part] #interval from R to tend
    log_encode$syspat[iclick] <- log_encode$Rtend_s[iclick] - log_encode$crop[iclick] # systolic ejection-period 
    
    # add 50ms security window (x) between systole end and start of diastole
    syspatx <- log_encode$syspat[iclick] + 0.05 # add security window of 0.05 (50ms) btw. systole and diastole
    Rtendx <- log_encode$Rtend_s[iclick] + 0.05
    
    # define diastole = RR - (syspatx + pep)
    log_encode$diaspat[iclick] <- log_encode$R_R_s[iclick] - (sitvec["pep"] + syspatx) 
    
    crop <- log_encode$crop[iclick] # pep - qonR
    syspat <- log_encode$syspat[iclick]  # Rtend - crop
    diaspat <- log_encode$diaspat[iclick] # remaining diastole
    
    # b) define in which intervals click falls
    # click in systole (bin syspat in two intervals sys1, sys2 of same size)
    log_encode$click_sys1[iclick] <-  as.numeric((dif <= (crop + 0.5*syspat)) & (dif > crop)) #sys1
    log_encode$click_sys2[iclick] <-  as.numeric((dif <= (crop + syspat)) & (dif > (crop + 0.5 * syspat))) #sys2

    # click in diastole
    log_encode$click_dias1[iclick] <- as.numeric((dif > Rtendx) & (dif <= (Rtendx + 0.25*diaspat)))
    log_encode$click_dias2[iclick] <- as.numeric(dif > (Rtendx + 0.25*diaspat) & (dif <= (Rtendx + 0.5*diaspat)))
    log_encode$click_dias3[iclick] <- as.numeric(dif > (Rtendx + 0.5*diaspat) & (dif <= (Rtendx + 0.75*diaspat)))
    log_encode$click_dias4[iclick] <- as.numeric(dif > (Rtendx + 0.75*diaspat) & (dif <= Rtendx + diaspat))
    
    # click in non-defined
    log_encode$click_pep[iclick] <- as.numeric((dif <= crop) | (dif > (Rtendx + diaspat))) #pep
    log_encode$click_x[iclick] <-  as.numeric((dif <= Rtendx) & (dif > (crop + syspat))) # security window
    
    # click_bin
    if (log_encode$click_pep[iclick] == 1) {
    log_encode$click_bin[iclick] <- "click_pep"
  } else if (log_encode$click_sys1[iclick] == 1| log_encode$click_sys2[iclick] == 1) {
    log_encode$click_bin[iclick] <- "click_lvet"
  } else if (log_encode$click_dias1[iclick] == 1| log_encode$click_dias2[iclick] == 1 |  log_encode$click_dias3[iclick] == 1 |  log_encode$click_dias4[iclick] == 1) {
    log_encode$click_bin[iclick] <- "click_dias"
  } else {
    log_encode$click_bin[iclick] <- "click_x"}
    
    # column total electromechanical systole (pep + ejection-period)
  log_encode$allsys[iclick] <- log_encode$pep[iclick] + log_encode$syspat[iclick] # allsys = pep + syspat
  log_encode$alldias[iclick] <- log_encode$diaspat[iclick] # alldias = diaspat
  
    }
}

#--------------------------------------------------------------------------
```


### 1.B.3. Define cardiac phases of hits, misses

* Analysis recognition (hit, miss), relative to phase of stimulus onset in encoding 
```{r add recognition parameters to log_encode}

# create new columns to check det = HIT, MISS relative to CLICK: sys vs. dias, dias_bin
log_encode$hit_pep <- as.numeric((log_encode$detection == "HIT") & (log_encode$click_pep == 1)) # hits in pep
log_encode$hit_sys1 <- as.numeric((log_encode$detection == "HIT") & (log_encode$click_sys1 == 1)) # hits in sys
log_encode$hit_sys2 <- as.numeric((log_encode$detection == "HIT") & (log_encode$click_sys2 == 1)) # hits in sys
log_encode$hit_x <- as.numeric((log_encode$detection == "HIT") & (log_encode$click_x == 1)) # hits in x

log_encode$hit_dias1 <- as.numeric((log_encode$detection == "HIT") & (log_encode$click_dias1 == 1)) # hits in dias
log_encode$hit_dias2 <- as.numeric((log_encode$detection == "HIT") & (log_encode$click_dias2 == 1)) # hits in dias
log_encode$hit_dias3 <- as.numeric((log_encode$detection == "HIT") & (log_encode$click_dias3 == 1)) # hits in dias
log_encode$hit_dias4 <- as.numeric((log_encode$detection == "HIT") & (log_encode$click_dias4 == 1)) # hits in dias

log_encode$miss_pep <- as.numeric((log_encode$detection == "MISS") & (log_encode$click_pep == 1))# miss in pep
log_encode$miss_sys1 <- as.numeric((log_encode$detection == "MISS") & (log_encode$click_sys1 == 1)) # miss in sys
log_encode$miss_sys2 <- as.numeric((log_encode$detection == "MISS") & (log_encode$click_sys2 == 1)) # miss in sys
log_encode$miss_x <- as.numeric((log_encode$detection == "MISS") & (log_encode$click_x == 1)) # miss in x

log_encode$miss_dias1 <- as.numeric((log_encode$detection == "MISS") & (log_encode$click_dias1 == 1)) # miss in dias
log_encode$miss_dias2 <- as.numeric((log_encode$detection == "MISS") & (log_encode$click_dias2 == 1)) # miss in dias
log_encode$miss_dias3 <- as.numeric((log_encode$detection == "MISS") & (log_encode$click_dias3 == 1)) # miss in dias
log_encode$miss_dias4 <- as.numeric((log_encode$detection == "MISS") & (log_encode$click_dias4 == 1)) # miss in dias

#--------------------------------------------------------------------------
```



## 1.C. Add additional variables to log_encode

### 1.C.1. Add ccovariates 

```{r add Covariates to log_encode}
#--------------------------------------------------------------------------
# add col covariates (IA, mean RR, HRV, STAI) to dataframe log_encode
log_encode <- left_join(log_encode, subset(score, select = c(IA,vp)), by = c('vp')) # add IA
log_encode <- left_join(log_encode, subset(kubios, select = c(mrr_rest, mhr_rest, rmssd_rest,vp)), by = c('vp')) # add rmssd from rest
log_encode <- left_join(log_encode, subset(stai, select = c(stai_sum,vp)), by = c('vp')) # add STAI

log_encode <- left_join(log_encode, subset(data_age, select = c(age,vp)), by = c('vp')) # add age
log_encode <- left_join(log_encode, subset(kubios_main, select = c(mhr_main,vp)), by = c('vp')) # add mean heart rate from experimental period

#--------------------------------------------------------------------------
```


### 1.C.2. Add ratings 

* Add individual and normative ratings for each picture to log_encode
```{r add ratings to log_encode}
#--------------------------------------------------------------------------
# subset individual ratings from log_rate -> join with log_encode
log_rate_subs <- log_rate[,c("vp", "stimnum", "rate_valence", "rate_arousal")] # subset with individual ratings
log_encode$rate_arousal <- NULL
log_encode <- left_join(log_encode, log_rate_subs, by = c('vp', 'stimnum'))

# add normative arousal ratings to log_encode
log_encode <- left_join(log_encode, emopics_norms[1:3], by = c('stimnum'))

# save(log_encode, file = paste(path_dataframes,"log_encode.RData", sep="")) # save log_encode
#--------------------------------------------------------------------------
```




# 2. PREPARE SHORT DATA FRAME: DATA_BINS

## 2.A. Create data_bins (all valences)
* Create short df with one row per pearticipant
* melted -> casted log_encode: one row per vp
```{r create df data_bins}
#--------------------------------------------------------------------------
# create data_bins -> cast means
m <- melt(log_encode, id= c("vp", "valence"), measure = c("dur_enc_min", "HR_1perMin", "R_R_s", "Rtend_s", "qonR", "crop", "pep", "syspat", "diaspat", "allsys", "alldias"))
data_bins <- cast(m, formula = vp ~ variable, mean) # c(mean)

#--------------------------------------------------------------------------
# sum of clicks in each cardiac interval
m1 <- melt(log_encode, id = c("vp", "valence"), measure = c("click_pep", "click_sys1", "click_sys2", "click_x", "click_dias1", "click_dias2", "click_dias3", "click_dias4")) # long df of 0,1 
c1 <- cast(m1, formula = vp ~ variable, sum) # cast -> calculate sum of clicks for each vp
data_bins <- left_join(data_bins, c1, by = c("vp"))

#--------------------------------------------------------------------------
# sum of detection parameters (hits, misses, fa, cr) in each cardiac interval
m2 <- melt(log_encode, id = c("vp", "valence"), measure = c("hit_pep", "hit_sys1", "hit_sys2", "miss_pep", "miss_sys1", "miss_sys2", "hit_dias1", "hit_dias2", "hit_dias3", "hit_dias4", "miss_dias1", "miss_dias2", "miss_dias3", "miss_dias4")) # long df of 0,1
c2 <- cast(m2, formula = vp ~ variable, sum) # cast -> calculate sum
data_bins <- left_join(data_bins, c2, by = c("vp"))

#--------------------------------------------------------------------------
# define proportion of heart intervals relative to total cardiac cycle
data_bins$prop_allsys_RR <- data_bins$allsys / data_bins$R_R_s
data_bins$prop_syspat_RR <- data_bins$syspat / data_bins$R_R_s # proportion syspat (=ejection period)
data_bins$prop_pep_RR <- data_bins$pep / data_bins$R_R_s # proportion pep (=pre-ejection period)
data_bins$prop_alldias_RR <- data_bins$alldias / data_bins$R_R_s

#--------------------------------------------------------------------------
# count sum of clicks in total electromechanical systole (pep+sys1+sys2), ejection period (lvet = sys1+sys2), diastole (dias1+dias2+dias3+dias4)
# check click_allsys + click_alldias + data_bins$click_x == 120
data_bins$click_allsys <-  data_bins$click_pep + data_bins$click_sys1 + data_bins$click_sys2 # clicks allsys 
data_bins$click_lvet <- data_bins$click_sys1 + data_bins$click_sys2 # clicks lvet 
data_bins$click_alldias <- data_bins$click_dias1 + data_bins$click_dias2 + data_bins$click_dias3 + data_bins$click_dias4 # clicks dias
```


## 2.B. Add Covariates (IA, STAI, HRV) to data_bins
```{r add covariates to data_bins}
#--------------------------------------------------------------------------
data_bins <- merge(data_bins, data_var, by = "vp") # join with data_bins

# save(data_bins, file = paste(path_dataframes,"data_bins.RData", sep="")) # save data_bins
#--------------------------------------------------------------------------
```

