---
title: "0303_INCASI_analysis"
author: "Stella"
date: "03. March 2018"
output: html_document
---

MAIN ANALYSIS:

**1. ENCODING**

1a. Circular analysis

1b. Binary analysis


**2. RECOGNITION**

2a. Circular analysis

2b. Binary analysis


SUPPLEMENTARY ANALYSIS

1. Supplementary recognition memory models

2. Analyse individual picture ratings

3. Analayse recognition rates


### INCASI SETUP

#### 1. General setup
```{r INCASI setup, message = FALSE}
#--------------------------------------------------------------------------
# Clear workspace
rm(list=ls())

#--------------------------------------------------------------------------
# Set working directory to containing folder ('0303_INCASI')
setwd('..')
cat('Current WD is: ', getwd())

#--------------------------------------------------------------------------
# Define paths
path_wd 	 <- getwd()
path_data <- paste("../0303_INCASI_data/",sep="") # path to file: _data
path_scripts  <- paste(path_wd,"/_scripts/",sep="") # path to file: _scripts
path_functions  <- paste(path_wd,"/_functions/",sep="") # path to file: _functions
path_dataframes <- paste(path_wd,"/_dataframes/",sep="") # path to file: _dataframes
path_variables <- paste(path_wd,"/_variables/",sep="") # path to file: _variables
path_figures <- paste(path_wd,"/_figures/",sep="") # path to file: _figures

#--------------------------------------------------------------------------
# load packages
source(paste(path_scripts,"loadPackages.R",sep="")) # installed Rpackages

#--------------------------------------------------------------------------
# load settings for ggplot
source(paste(path_scripts,"loadTheme.R",sep="")) # ggplot theme
source(paste(path_scripts,"loadColors.R",sep="")) # ggplot colours

#--------------------------------------------------------------------------
# define list of participants
inc_list <- as.factor(dir(path_data)) #create list with participatns with all inc04 - inc50

#--------------------------------------------------------------------------
```


####  2. Load dataframes for analysis: 

Coding of cardiac phases:
* Systole (lvet = syspat): sys1 + sys2
* Diastole (alldias): dias1-dias4 (4 diastolic intervals of same length)
* Non-defined: pep (pre-ejection phase), security window (50ms between end of systole and start diastole)

* Load dataframes created in preprocessing scripts
```{r scripts analysis, message = FALSE}
#----------------------------------------------------------------------------
# dataframes for analysis behaviour relative to cardiac cycle
load(paste(path_dataframes,"log_encode.RData", sep="")) # long dataframe encoding (all datapoints from encoding)
load(paste(path_dataframes,"data_bins.RData", sep="")) # short dataframe encoding (1 subject each row)

# dataframes for analysis rating
load(paste(path_dataframes, "log_rate.RData", sep="")) # long dataframe rating (all datapoints from rating period)

load(paste(path_dataframes, "rate_va.RData", sep="")) # df with normative and individual ratings for each picture (180 pictures)
#----------------------------------------------------------------------------
```


####  3. Load dataframes for covariates
```{r load covariate scripts, results='hide'}
#--------------------------------------------------------------------------
# Interoceptive Awareness (IA)
load(paste(path_dataframes, "score.RData", sep="")) # heart beat perception score df

# Heart rate variability (HRV)
load(paste(path_dataframes, "kubios.RData", sep="")) # kubios (mean HR, rmssd) df

# State Trait Anxiety (STAI)
load(paste(path_dataframes, "stai.RData", sep="")) # STAI-T (trait anxiety) df
#--------------------------------------------------------------------------
```


####  4. Load functions
```{r circular functions}
#----------------------------------------------------------------------------
# load functions for circular analysis
source(paste(path_functions, "fx_circ_encoding.R", sep="")) # circular function for picture onsets relative to cardiac cycle (encoding period)
source(paste(path_functions, "fx_circ_recognition.R", sep="")) # circular function for memory relative to picture onset in encoding (recognition period)

# load helper functions
source(paste(path_functions, "fx_standard_error.R", sep="")) # function to calculate standard error
#----------------------------------------------------------------------------
```


####  5. Clean participant list 
```{r inc_clean, results='hide'}
#--------------------------------------------------------------------------
# load clean inc_lists
source(paste(path_scripts, "0303_Clean_inc_list.R", sep ="")) # load clean participant list

# inc_clean <- inc_clean[c(1:40)] # take only fist 40 participants out of clean inc list (to test with dataset of preregistration size)
#--------------------------------------------------------------------------
```



# MAIN ANALYSIS

## 1. ENCODING - Cardiac influence on visual sampling

* Adapt dataframe to clean participant list
```{r enc adapt dataframe}
#--------------------------------------------------------------------------
# data_bins_enc: Create subset data_bins for encoding
log_encode <- log_encode[log_encode$vp %in% inc_clean,]
data_bins_enc <- data_bins[data_bins$vp %in% inc_clean,] 
#--------------------------------------------------------------------------
```


### 1.A. Circular analysis

* Relative onset of each self-chosen key press (i.e. stimulus onset) was computed across the cardiac cycle (from Rpeak to Rpeak), with radian measures (between 0 and 2π) assigned to each stimulus onset (c.f. `0303_INCASI_preprocess_b`)
* Distribution of stimulus onsets is tested against the circular unifrom with Rayleigh tests
* Apply function circ_click (loaded script *fx_circ_encoding.R*)

#### 1.A.1. Exemplary participant-level analysis
* Exemplary stimulus distribution (120 black dots) across the cardiac cycle (Rpeak = 0) of one participant

**Figure 2.b.**
* Orange segment: individual systole (start: solid, end: dashed)
* Blue segment: individual diastole (start: solid, end: dashed)
* Grey arrow: overall picture onset mean
* Grey line: circular density of picture onsets
```{r test participant-level against uniform}
#--------------------------------------------------------------------------
# plot exemplary individual plot  
# pdf(file = paste(path_figures, "circ_click_sub25.pdf", sep=""), width=5, height=5)
circ_click("inc25", plot1 = T, ray1 = T) # distribution of key-presses (picture onsets)
# dev.off()
#--------------------------------------------------------------------------
```


#### 1.A.2. Group-level analysis
* Distribution of participants' stimulus onset means (43 black dots) across the cardiac cycle
* Hypothesis: significant deviation of means from circular uniform distribution in early cardiac phase

Figure (for visualization purposes, not included in manuscript):
* Orange lines: individual means that fall into individual systole
* Blue lines: individual diastolic means
* Grey lines: individual means in non-defined cardiac intervals
* Black arrow: overall picture onset mean across participants
```{r test group-level against uniform}
#--------------------------------------------------------------------------
# 1. distribution of clicks in encoding
par(mar=c(0,0,0,0))
# pdf(file = paste(path_figures, "circ_click.pdf", sep=""),width=5,height=5)
circ_click(inc_clean, plot2 = T, ray2 = T) # distribution of clicks without inc_HyperHR, inc_Err
# dev.off()

# print overall circular mean and sd (in radian)
mean2 <- circ_click(inc_clean, mean2 = T)

# calculate circular mean in pi: mean2 = mean2pi *pi
mean2pi <- mean2[1] / pi
sd2pi <- mean2[4] / pi
#--------------------------------------------------------------------------
```


#### 1.A.3. Bootstrapping analysis
* Non-parametical computation of confidence intervals and significance (based on Ohl et al., 2016)

Bootstrap procedure:
* From the original paricipant pool (N=43), a bootstrap sample (N=43) is drawn (with replacement)
* For each participant in the bootstrap sample, a circular density (bw = 20) of picture onsets is computed
* In the next step, the mean circular density across all participants in the bootstrap sample is computed

* This bootstrap procedure is repeated 10000 times
* 95% confidence intervals are determined as 2.5% and 97.5% percentiles from the distribution of mean circular densities
```{r boostrap}
#--------------------------------------------------------------------------
# bootstrap main analysis
nBoot  <- 10000 # number of repition of bootstrap procedure
vplist <- unique(log_encode$vp) # list of participants

# set up output variables
out    <- c()
buffer <- c()
bwparam <- 20 # bandwith parameter

# a) run whole bootstrap procedure
# start loop
for(i in 1:nBoot){ # repeat bootstrap procedure nBoot times
  tmme <- system.time({
  vplist_boot <- sample(vplist,replace=TRUE) # draw bootstrap sample
  buffer <- c()
  for(j in 1:length(vplist_boot)){ # compute circular density for each participant in sample
    x=circular(log_encode$radclick[which(log_encode$vp==vplist_boot[j])],type="angles",units="radians",modulo="2pi",zero=pi/2,rotation="clock")
    res50x  <- density(x, bw=bwparam) #Play around with bw. Make chi-square tests and check whether observed coupling is observed for all kinds of bw parameters
    readout <- as.vector(res50x$y)
    buffer  <- rbind(buffer,readout,deparse.level=0)
  }
  buffer  <- colMeans(buffer)
  out <- rbind(out,buffer)})
  if (i%%100 == 0) {
    print(paste('Iteration ', num2str(i, fmt=0), ' of ', num2str(nBoot, fmt=0)))
    print(paste('Time until end: ', num2str(tmme[3]*(nBoot-i)/60), 'minutes'))
  }
}

# save(res50x, file = paste(path_dataframes,"res50x.RData", sep="")) # save density
# save(out, file = paste(path_dataframes,"out.RData", sep="")) # save bootstrap output

#--------------------------------------------------------------------------
# b) if boostrap procedure is already run, load bootstrap dataframes
# load bootstrap dataframes
# load(paste(path_dataframes,"res50x.RData", sep="")) 
# load(paste(path_dataframes,"out.RData", sep="")) 

# get 95% confidence intervals and median from bootstraps (quantiles: 0.025, 0.5, 0.975)
ci_upper  <- c()
ci_lower  <- c()
ci_median <- c() 
for(i in 1:ncol(out)){
  ci_lower  <- c(ci_lower, as.numeric(quantile(out[,i],0.025))) # 2.5% percentile 
  ci_upper  <- c(ci_upper, as.numeric(quantile(out[,i],0.975))) # 97.5% percentile
  ci_median <- c(ci_median, as.numeric(quantile(out[,i],0.5))) # median
}
markup_upper <- which(ci_lower>dcircularuniform(1)) # upper CI bound
markup_lower <- which(ci_upper<dcircularuniform(1)) # lower CI bound
#--------------------------------------------------------------------------
```


Plot bootstrap density on circular plot with individual picture onset means:

**Figure 1.a**:
* Circular distribution of individual mean picture onsets (black dots, N = 43)
* Black arrow: weighted overall mean of picture onsets as black arrow
* Middle thicker line: mean circular density of picture onsets (based on bootstrap procedure)
* Orange segment: mean systole
* Blue segment: mean diastole
* Grey: non-defined phases
* Inner and outer thin grey lines: inner and out bound of 95% CI
```{r boostrap plot}
#--------------------------------------------------------------------------
# plot bootstrap results

# 1. load circular data of 2nd level analysis (circular mean picture onsets for each participant)
load(paste(path_dataframes,"H_rad_secondlevel.RData", sep="")) 

par(mar=c(0,0,0,0))
# pdf(file = paste(path_figures, "bootstrap_bw",bwparam,".pdf",sep=""),width=6.5,height=6.5)

# 2. plot circular plot with mean picture onsets (43 black dots)
plot(H_rad_secondlevel, stack=TRUE, bins = 720, shrink = 1.0, axes = T, col= "black",cex = 1.0, lwd = 3, 
     xlim=c(-1.22,1.22),ylim=c(-1.22,1.221)) # circular plot of 2nd level means

# 3. add bootstrap density to plot
res50x$y <- ci_median/dcircularuniform(1) 
lines.circular(res50x$x,res50x$y,join=TRUE,col=defgrey,offset=0,rotation="clock",lwd=4)

# 4. compute mean length of cardiac phases (systole, diastole) over all participants
# transform mean lengths of cardiac phases into radian
dat <- data_bins_enc
mcirctrans <- mean(2 * pi * 1/(dat$R_R_s)) # circular transformation
msysstart <- mean(mcirctrans * (dat$crop)) # mean systole start
msysend <- mean(mcirctrans * (dat$Rtend_s)) # mean systole end
mdiasstart <- mean(mcirctrans * (dat$Rtend_s + 0.05)) # mean diastole start
mdiasend <- mean(mcirctrans * (dat$Rtend_s + 0.05 + dat$diaspat)) # mean diastole end

# 5. colour the systolic and diastolic segments of the circular density
x <- res50x$x # x coordinates of points on density
msyspos <- which(x >= msysstart & x <= msysend) # position of points for systole
mdiaspos <- which(x >= mdiasstart & x <= mdiasend) # position of points for diastole

lines.circular(res50x$x[msyspos],res50x$y[msyspos],join=F,col=deforange,offset=0,rotation="clock",lwd=3) # draw systolic density segment (orange)
lines.circular(res50x$x[mdiaspos],res50x$y[mdiaspos],join=F,col=defmedblue,offset=0,rotation="clock",lwd=3) # draw diastolic density segment (blue)

# #check fit of segments (compute beginning and end of systolic, diastolic segment)
# # function to draw line segments onto circular plot
# circseg <- function (a, lty, col) { # define a= radians value to draw segment at, lty = linetype (1 = normal, 2 = dashed)
#   xcoord <- sin(a)
#   ycoord <- cos(a)
#   segments(0,0,xcoord, ycoord, col = col, lty= lty, lwd = 2)
# }
# 
# # check fit of density segments
# circseg(msysstart, 1, deforange) # sys start
# circseg(msysend, 2, deforange) # sys end
# circseg(mdiasstart, 1, defmedblue) # dias start
# circseg(mdiasend, 2, defmedblue) # dias stop

# 6. plot significant density segment
lines.circular(res50x$x[markup_upper],res50x$y[markup_upper],col=deforange,lwd=8,offset=0,rotation="clock")

# 7. plot lower and upper bound of confidence interval
res50x$y <- ci_lower/dcircularuniform(1) # lower bound
lines.circular(res50x$x,res50x$y,join=TRUE,col="light grey",offset=0,rotation="clock",lwd=2)
res50x$y <- ci_upper/dcircularuniform(1) # upper bound
lines.circular(res50x$x,res50x$y,join=TRUE,col="light grey",offset=0,rotation="clock",lwd=2)

# 8. plot central cross to mark center of circle
segments(0,0,0.05,0, lwd = 2)
segments(0,0,-0.05,0, lwd = 2)
segments(0,0.05,0,0, lwd = 2)
segments(0,-0.05,0,0, lwd = 2)

# 9. plot overall circular mean of picture onsets
arrows.circular(mean(H_rad_secondlevel), y=rho.circular(H_rad_secondlevel), lwd = 5, col = "black") # overall mean

# dev.off()
#--------------------------------------------------------------------------
```


### 1.B. Binary analysis

Analysis of individual cardiac phases: compare phases (individual systole, diastole) against each other 

Analyse ratios between both carduac phases (systole, diastole) to take into account inter-individual differences in phase lengths:
* Sum of clicks (picture onsets) per phase (as ratio of all 120 trials) is normalized to the proportion of the subject-specific phase length in the total cardiac cycle
* Define ratios: (clicks per phase / 120) / (individual phase length/ individual mean R-R length)

* With no cardiac effect: clicks (triggering picture onsets) would expected to be randomly distributed across both cardiac phases -> click ratio (clicks/120) should correspond to cardiac phase ratio (phase length/whole cycle) -> ratio value = 1
* With cardiac effect: over-proportional accumulation of key presses in systole -> systolic ratio value > 1

* Hypothesis: significant accumulation of clicks in individual systole: ratio value for systole > diastolic ratio


#### 1.B.1 Define ratios for both phases (systole, diastole):
```{r enc define ratios}
#--------------------------------------------------------------------------
# compute ratios for each phase

## SYSTOLE
# each systolic phase
data_bins_enc$click_pep_rel <- (data_bins_enc$click_pep/120) /  (data_bins_enc$pep/data_bins_enc$R_R_s) # pre-ejection period
data_bins_enc$click_sys1_rel <- (data_bins_enc$click_sys1/120) /  ((0.5 * data_bins_enc$syspat)/data_bins_enc$R_R_s) # first 1/2 of ejection-phase 
data_bins_enc$click_sys2_rel <- (data_bins_enc$click_sys2/120) / ((0.5 * data_bins_enc$syspat)/data_bins_enc$R_R_s) # second 1/2 of ejection-phase

# whole ejection-phase
data_bins_enc$click_lvet_rel <- ((data_bins_enc$click_lvet)/120)/((data_bins_enc$syspat)/data_bins_enc$R_R_s) # all clicks in lvet (sys1+sys2) / lvetlength (syspat) 

# total electromechanical systole (pre-ejection + ejection phase)
data_bins_enc$click_allsys_rel <- ((data_bins_enc$click_allsys)/120)/((data_bins_enc$pep + data_bins_enc$syspat)/data_bins_enc$R_R_s)  # all clicks in sys (pep+sys1+sys2) / syslength (pep + syspat)

#--------------------------------------------------------------------------
## DIASTOLE
# each diastolic phase (whole diastole divided in 4 parts)
data_bins_enc$click_dias1_rel <- (data_bins_enc$click_dias1/120) / ((0.25* data_bins_enc$diaspat)/data_bins_enc$R_R_s) 
data_bins_enc$click_dias2_rel <- (data_bins_enc$click_dias2/120) / ((0.25* data_bins_enc$diaspat)/data_bins_enc$R_R_s) 
data_bins_enc$click_dias3_rel <- (data_bins_enc$click_dias3/120) / ((0.25* data_bins_enc$diaspat)/data_bins_enc$R_R_s) 
data_bins_enc$click_dias4_rel <- (data_bins_enc$click_dias4/120) / ((0.25* data_bins_enc$diaspat)/data_bins_enc$R_R_s)  

# total diastole
data_bins_enc$click_alldias_rel <- ((data_bins_enc$click_alldias)/120) / (data_bins_enc$diaspat/data_bins_enc$R_R_s) # all clicks in dias (dias1+dias2+dias3+dias4) / diaslength (diaspat)
```


#### 1.B.2 Run analysis: Paired t-test
* Group-level analysis: normalized systolic and diastolic ratios are tested against each (two-sided paired t-test)
* Is there a significant higher systolic (lvet) vs. diastolic ratio?
```{r enc paired ttest}
#--------------------------------------------------------------------------
# paired t-test
t.test(data_bins_enc$click_lvet_rel, data_bins_enc$click_alldias_rel, paired = T) # *
cohen.d(data_bins_enc$click_lvet_rel, data_bins_enc$click_alldias_rel, paired = T)
```


#### 1.B.3. Plot results with ggplot

a) Prepare dataframes: melt and cast dataframes
* phase (lvet, dias = variable) ~ respective ratios (=value)
```{r enc  prep dataframes}
#--------------------------------------------------------------------------
# prepare long df m_lvetclick (relative clickrates for ejection-phase vs whole diastole)
# melt data_bins_enc showing phase (lvet, dias = variable) ~ respective ratios (=value)
m_lvetclick <- melt(data_bins_enc, id = "vp", measure = c("click_lvet_rel", "click_alldias_rel"))
colnames(m_lvetclick) <- c("vp", "phase", "rel_clickrate")

# cast m_lvetclick: for each phase(lvet, dias=variable) show ratio(=value) mean, sd, length, se
c_lvetclick <- cast(m_lvetclick, phase ~ ., c(mean, sd, length, se, se_up, se_down), value = "rel_clickrate")

#--------------------------------------------------------------------------
```


b) Create column for colouring different cardiac phases in ggplot (systole, diastole, non-defined phase)
```{r colour column}
#--------------------------------------------------------------------------
## COMPARE SYSTOLE(LVET) AND DIASTOLE

# compare click rates with phase rates (click rate - phase rate)
# systole
data_bins_enc$diff_plvet <- (data_bins_enc$click_lvet/120) - data_bins_enc$prop_syspat_RR # click rate systole - prop systole
data_bins_enc$check_plvet <- (sign(data_bins_enc$diff_plvet)) # check whether click rate (1) or phase rate (-1) is higher

# diastole
data_bins_enc$diff_pdias <- (data_bins_enc$click_alldias/120) - data_bins_enc$prop_alldias_RR # click rate dias - prop dias
data_bins_enc$check_pdias <- (sign(data_bins_enc$diff_pdias))

# create column col to prepare coloring each quadrant (q1, q2, q3,q4) of ggplot
for (i in (1:length(data_bins_enc$vp))) {
  if(as.numeric(data_bins_enc$check_pdias[i] > 0 & data_bins_enc$check_plvet[i] < 0)) { #q1 (left upper, overproportional diastolic ratio)
    data_bins_enc$col[i] <- 1
  } else if(data_bins_enc$check_pdias[i] > 0 & data_bins_enc$check_plvet[i] > 0) { #q2 (right upper, overproportional diastolic ratio)
    data_bins_enc$col[i] <- 2
  } else if (data_bins_enc$check_pdias[i] < 0 & data_bins_enc$check_plvet[i] > 0) { #q3 (right lower, overproportional systolic ratio)
    data_bins_enc$col[i] <- 3
  } else {
    data_bins_enc$col[i] <- 2 #q4 (left lower, non-defined)
  }
  
}
data_bins_enc$col <- factor(data_bins_enc$col)
#--------------------------------------------------------------------------
```


c) Scatter plot for systolic vs. diastolic ratio
**Figure 1.b**:
* Systolic vs. diastolic ratio for each participant
* Dashed lines: number of picture onsets that would be expected if they were uniformly distributed (resulting in a ratio = 1)
* Orange: participants who preferred to prompt pictures in systole (systolic ratio >1, diastolic <1)
* Blue: participants who preferred to prompt pictures in diastole (diastolic ratio >1, systolic <1)
* Grey: participants who did not show a preference in any of the two defined phases
```{r ggplot relative ratios}
#--------------------------------------------------------------------------
par(mar=c(0,0,0,0))
# pdf(file = paste(path_figures, "fig_sysdias1",".pdf",sep=""),width=4.5,height=4.5)

# scatterplot relative click rates for total systole (x) vs. total diastole (y)
fig_sysdias <- ggplot(data_bins_enc) +
  geom_hline(aes(yintercept = 1), size=0.5, linetype="dashed")+ 
  geom_vline(aes(xintercept = 1), size=0.5, linetype="dashed")+
  geom_point(size=3, aes(x=click_lvet_rel, y=click_alldias_rel, col = col, shape = col))+ #click_alldias_rel
  scale_colour_manual(values=c(defmedblue, defgrey, deforange, defgrey)) +
  geom_point(data = c_lvetclick, colour = "black", aes(x=mean[1], y=mean[2]), size = 3) +
  geom_errorbarh(data = c_lvetclick, colour = "black", aes(x = mean[1], xmin = se_down[1], xmax = se_up[1], y = mean[2]),  size=0.5, height = 0.005) +
  geom_errorbar(data = c_lvetclick, colour = "black", aes(ymin = se_down[2], ymax = se_up[2], x = mean[1]),  size=0.5, width = 0.005) +
  coord_cartesian(xlim = c(0.75, 1.4), ylim = c(0.75, 1.4)) +
  labs(x = "Systolic ratio", y = "Diastolic ratio") +
  mytheme 

fig_sysdias

# dev.off()

#--------------------------------------------------------------------------
```




## 2. RECOGNITION - Cardiac influence on recognition memory

* Adapt dataframe to clean participant list
```{r rec analysis absolute}
#--------------------------------------------------------------------------
log_encode <- log_encode[log_encode$vp %in% inc_clean,]
#--------------------------------------------------------------------------
```


### 2.A. Circular analysis

* Test distribution of memory probes (hits, misses) relative to their cardiac time point in encoding against circular uniform distribution (for each valence: neg, neu, pos)


#### 2.A.1. Exemplary participant-level analysis of hits, misses

* Exemplary stimulus distribution of picture onsets across the cardiac cycle (Rpeak = 0) for hits, misses
* Circular plots (for visualization purposes, not included in manuscript):
```{r test HIT, MISS against uniform for one vp}
#--------------------------------------------------------------------------
# first-level analysis: exemplary participant
circ_click_mem("inc25", det = "hit_miss", plot1 =T, ray1  = T) # distribution of memory probes (= hits and misses)
circ_click_mem("inc25", det = "HIT", plot1 =T, ray1  = T) # distribution of hits
circ_click_mem("inc25", det = "MISS", plot1 =T, ray1  = T) # distribution of misses
#--------------------------------------------------------------------------
```


#### 2.A.2. Group-level analysis of hits, misses 

* Distribution of participants' stimulus onset means (encoding) across the cardiac cycle for hits and misses
* Circular plots (for visualization purposes, not included in manuscript):
```{r test HIT, MISS against uniform group-level}
#--------------------------------------------------------------------------
# second-level analysis: HITs
# pdf(file = paste(path_figures, ,"circ_det_hit.pdf", sep=""), width=6, height=6)
circ_click_mem(inc_clean_cutSignmem_ray, det = "HIT", plot2 =T, ray2 = T) 
# dev.off()

# second-level analysis: MISSes
# pdf(file = paste(path_figures, "Manuscript/","circ_det_miss.pdf", sep=""), width=6, height=6)
circ_click_mem(inc_clean_cutSignmem_ray, det = "MISS", plot2 =T, ray2 = T)
# dev.off()
#--------------------------------------------------------------------------
```



### 2.B. Binary analysis: Linear mixed model (LMM) 

* Analyse influence of cardiac phase (systole, diastole) and valence (negative, neutral, positive) on recognition memory.

* LMM for binomial data with subject as random factor
* Dependent variable: recognition memory (coding: 0 = miss, 1 = hit) = bivariate dependent variable

* Independent within-subject variables:
  * Cardiac phase: 0 = diastole, 1 = systole
  * Three valence levels (positive, negative, neutral), contrast-coded with neutral valence as baseline condition:   
    positive-neutral, negative-neutral


#### 2.B.1. Compute overall mean recognition performance

* Mean recognition performance (detection rate) for negative, neutral, positive old pictures 
```{r mean recognition performance}
#--------------------------------------------------------------------------
## adapt dataframe
data <- log_encode[log_encode$vp %in% inc_clean,] 

# remove all trials with items that were not presented during encoding (include only old pictures)
idx <- which(is.na(data$answer)==FALSE) # select only rows where answer is not NA (NA for new pictures)
d <- data[idx,] # df for memory probes 

#--------------------------------------------------------------------------
# mean performance for each participant overall
m1 <- melt(d,id=c("vp"),measure=c("answer"))
c1 <- cast(m1, vp ~ variable,mean)

# get mean recognition performance for each participant (for each valence)
m2 <- melt(d,id=c("valence","vp"),measure=c("answer"))
c2 <- cast(m2,valence + vp ~ variable,mean)

#mean recognition performance over all participants for each valence
m3 <- melt(c2,id=c("valence"),measure=c("answer"))
c3 <- cast(m3, valence ~ variable, c(mean, sd))
#--------------------------------------------------------------------------
```



#### 2.B.2. Prepare variables
* Valence: contrast-coded with neutral as baseline (positive-neutral, negative-neutral)
* Cardiac phase: 0 = diastole, 1 = systole
```{r regression model rec prep}
#--------------------------------------------------------------------------
## prepare variables for linear model
d$vp <- as.factor(d$vp)

# adapt dataframe: select only trials from individual systole, diastole
# phase: set diastole == 0 as reference condition, systole = 1
d <- d[(d$click_bin=="click_lvet" | d$click_bin=="click_dias"),] # select only clicks in systole (lvet) and diastole (dias)
d$ds <- as.factor(ifelse(d$click_bin=="click_dias",0,1)) # set dias 0, lvet 1 (factor with 2 levels)

# valence: define contrast - set neutral valence as reference condition
d$valence2 <- d$valence
contrasts(d$valence2) <- contr.treatment(3,base=1) # contrast with neutral as baseline -> 3 factor levels, base = neutral (level1)
#--------------------------------------------------------------------------
```


#### 2.B.3. Run model
* m0: recognition memory ~ valence
* m1: recognition memory ~ valence * cardiac phase

* see also **Table 1**:
```{r regression model rec run}
#--------------------------------------------------------------------------
# 1. model for valence
m0 <- glmer(answer~ valence2 + (1|vp),data=d,family="binomial") # valence

# 2. model for valence * cardiac phase (include arousal ratings? -> iar, iarz)
m1 <- glmer(answer ~ valence2*ds + (1|vp),data=d,family="binomial") # valence * cardiac phase

# check output of models
summary(m0) 
#--------------------------------------------------------------------------
```


#### 2.B.4. Plot model
* Plot model prediction (for visualization purposes, not included in manuscript):
```{r regression model rec plot}
#--------------------------------------------------------------------------
## 1. overall model for valence:
# m0: memory performance as function of valence (m0)
d$pred_logit <- predict(m0) # add model fits to df
d$pred_prob  <- exp(d$pred_logit)/(1+exp(d$pred_logit))

#val: plot(d$answer,d$pred_prob)
pred_val <- ggplot(d,aes(x=valence2,y=pred_prob))+ 
  scale_colour_manual(values=c(defmedblue, deforange)) +
  geom_boxplot()+
  labs(x = "Valence", y = "Predicted probability") +
  mytheme
pred_val

#--------------------------------------------------------------------------
## 2. create model for valence * cardiac phase
# m1: compute predictions for model valence * cardiac phase (m1, m2)
d$pred_logit <- predict(m1) # add model fits to df
d$pred_prob  <- exp(d$pred_logit)/(1+exp(d$pred_logit))

#valds: plot(d$answer,d$pred_prob)
pred_valds <- ggplot(d,aes(x=valence2,y=pred_prob,colour=click_bin))+
  scale_colour_manual(values=c(defmedblue, deforange)) + # blue: diastole, orange: systole
  geom_boxplot()+
  labs(x = "Valence * Cardiac phase", y = "Predicted probability") +
  mytheme
pred_valds
#--------------------------------------------------------------------------
```




# SUPPLEMENTARY ANALYSIS

## 1. Supplementary models for recognition memory: add inter-individual variables

### 1.A. Include additional variables (inter-individual differences)
* Model: Recognition memory ~ valence * phase + additional variable

* 1. Heart rate variability (rmssdl: log-transformed to mitigate skewedness and centred to the mean) 
* 2. Trait Anxiety (staiz: z-transformed)
* 3. Interoceptive Awareness (IAz: z-transformed)
```{r regression model rec var prep}
#--------------------------------------------------------------------------
# 1. HRV (rmssd): log-transform of heart rate variability
d$rmssd <- d$rmssd_rest
d$rmssdl <- log(d$rmssd) - mean(log(d$rmssd)) # centered by mean
# hist(d$rmssdl)

# 2. STAI: z-Transform of trait anxiety
d$staiz <- (d$stai_sum - mean(d$stai_sum))/sd(d$stai_sum) 
#hist(d$staiz)

# 3. IA: z-Transform of interoceptive awareness
# remove 2 sujbects without IA measure (d1)
idx <- which(is.na(d$IA)==FALSE)
d1 <- d[idx,]
d1$IAz <- (d1$IA - mean(d1$IA))/sd(d1$IA)

# save(d, file = paste(path_dataframes,"d.RData", sep=""))
#--------------------------------------------------------------------------
```


### 1.B. Run supplementary models with additional covariates

* m2: add rmssdl to m1 (recognition memory ~ valence * cardiac phase + rmssdl)
* m2a: add age to m2 (recognition memory ~ valence * cardiac phase + rmssdl + age)

* m3: add staiz to m1 (recognition memory ~ valence * cardiac phase + staiz)
* m4: add IAz to m1 (recognition memory ~ valence * cardiac phase + IAz)

* see also **Table 1**:
```{r regression model rec var run}
#--------------------------------------------------------------------------
## 1. HRV: rmssd (with age as confounding variable)

# add rmssdl to m1
m2 <- glmer(answer ~ valence2*ds + rmssdl + (1|vp),data=d, family="binomial")
m2a <- glmer(answer ~ valence2*ds + rmssdl + age + (1|vp),data=d,family="binomial")

anova(m1,m2) # *
anova(m1,m2a) # 

# for visualisation
gam2 <- bam(answer ~ valence2*ds + s(rmssdl),data=d,family="binomial")
plot(gam2)


#--------------------------------------------------------------------------
## 2. STAI
m3 <- glmer(answer ~ valence2*ds + staiz + (1|vp),data=d,family="binomial")

anova(m1,m3) # n.s.

#--------------------------------------------------------------------------
## 3. IA (run models with d1, minus 2 participants without IA)
# rerun m1 with d1 -> adapted participant number 
m1edit <- glmer(answer~valence2*ds + (1|vp),data=d1,family="binomial") # encoding durtaion
m4 <- glmer(answer~valence2*ds + IAz + (1|vp),data=d1,family="binomial") # encoding durtaion

anova(m1edit,m4) # n.s.

# output of model 
summary(m2) 
#--------------------------------------------------------------------------
```



## 2. Analyse ratings: Subjective perception of picture emotionality (normative vs. individual ratings)

### 2.A. Normative and individual means of picture ratings for arousal and valence

* Compare individual vs. normative ratings (mean, sd, se)

* see also **Supplementary Table 1**:
```{r stimnum ratings}
#----------------------------------------------------------------------------
# 1. AROUSAL: melt arousal ratings - melt rate_va
m_rate_a <- melt(rate_va, id = c("stimnum", "valence"), measure = c("rate_arousal", "norms_arousal"))

# cast m_rate_a: valence (rows) ~ mean individual rating (cols)
c_rate_a <- cast(m_rate_a, valence ~ variable, c(mean, sd, length, se)) #   mean, sd, length, se for each valence
c_rat_a_all <- cast(m_rate_a, . ~ variable, c(mean, sd, length, se)) # over all valences

# 2. VALENCE: melt valence ratings - melt rate_v
m_rate_v <- melt(rate_va, id = c("stimnum", "valence"), measure = c("rate_valence", "norms_valence"))

# cast m_rate_v: valence (rows) ~ mean individual rating (cols)
c_rate_v <- cast(m_rate_v, valence ~ variable, c(mean, sd, length, se)) # mean, sd, length, se for each valence
c_rat_v_all <- cast(m_rate_v, . ~ variable, c(mean, sd, length, se)) # over all valences
#----------------------------------------------------------------------------
```


### 2.B. Run tests to compare normative vs. individual ratings
* 1. Run ANOVA to test ratings across rating category (normative, individual) and valence
* 2. Run one-sided ttests for normative vs. individual ratings across each valence level

* see also **Supplementary Table 1**:
```{r stimnum ratings run tests}
#----------------------------------------------------------------------------
## 1. Run ANOVA:
# for each valence level: run anova to test for diff. in mean ratings between individual and normative 

# 1.1. AROUSAL ratings: rating value ~ valence * rating category (individual vs. norms)
aov_a <- aov((value) ~ valence * variable, data=m_rate_a) 
summary(aov_a)


# 1.2. VALENCE ratings
aov_v <- aov(log(value) ~ valence * variable, data=m_rate_v)
summary(aov_v)

#----------------------------------------------------------------------------
## 2. Run one-sided t-tests across valence levels

# run t-tests for each valence subset
# subsets for each valence
sub_neg <- rate_va[rate_va$valence == "negativ",]
sub_pos <- rate_va[rate_va$valence == "positiv",]
sub_neu <- rate_va[rate_va$valence == "neutral",]

## 2.1. AROUSAL
# a) Neutral pictures
a_neu <- t.test(sub_neu$norms_arousal,sub_neu$rate_arousal) # **
cohen.d(sub_neu$norms_arousal,sub_neu$rate_arousal)
cor.test(sub_neu$norms_arousal,sub_neu$rate_arousal) # ***

# b) Positive pictures
a_pos <- t.test(sub_pos$norms_arousal,sub_pos$rate_arousal) # ***
cohen.d(sub_pos$norms_arousal,sub_pos$rate_arousal)
cor.test(sub_pos$norms_arousal,sub_pos$rate_arousal) # ***

# c) Negative pictures
a_neg <- t.test(sub_neg$norms_arousal,sub_neg$rate_arousal) # *
cohen.d(sub_neg$norms_arousal,sub_neg$rate_arousal)
cor.test(sub_neg$norms_arousal,sub_neg$rate_arousal) # ***

# adjust p-values
p_aro <- c(a_neu$p.value, a_pos$p.value, a_neg$p.value)
p.adjust(p_aro, method = "bonferroni")

#----------------------------------------------------------------------------
## 2.2. VALENCE
# a) Neutral pictures
v_neu <- t.test(sub_neu$norms_valence,sub_neu$rate_valence) # n.s.
cohen.d(sub_neu$norms_valence,sub_neu$rate_valence)
cor.test(sub_neu$norms_valence,sub_neu$rate_valence) # ***

# b) Positive pictures
v_pos <- t.test(sub_pos$norms_valence,sub_pos$rate_valence) # ***
cohen.d(sub_pos$norms_valence,sub_pos$rate_valence)
cor.test(sub_pos$norms_valence,sub_pos$rate_valence) # ***

# c) Negative pictures
v_neg <- t.test(sub_neg$norms_valence,sub_neg$rate_valence) # n.s.
cohen.d(sub_neg$norms_valence,sub_neg$rate_valence)
cor.test(sub_neg$norms_valence,sub_neg$rate_valence) # ***

# adjust p-values
p_val <- c(v_neu$p.value, v_pos$p.value, v_neg$p.value)
p.adjust(p_val, method = "bonferroni")

#----------------------------------------------------------------------------
## 3. compute difference in norms arousal ratings: btw. negative and positive pictures (described in methods)
# norms ratings
t.test(sub_neg$norms_arousal,sub_pos$norms_arousal) # n.s. difference btw norms arousal (neg vs pos)
cohen.d(sub_neg$norms_arousal,sub_pos$norms_arousal)

# individual ratings
t.test(sub_neg$rate_arousal,sub_pos$rate_arousal) # *** diff. btw. rated arousal (neg vs pos)
cohen.d(sub_neg$rate_arousal,sub_pos$rate_arousal)
#----------------------------------------------------------------------------
```


### 2.C. Plot normative vs. individual ratings

* 1. Plot Valence ratings
* 2. Plot Arousal ratings
* see **Supplementary Figure 1**:
```{r stimnum ratings plot}
#----------------------------------------------------------------------------
# 1. Valence ratings
cor_val <- ggplot(rate_va) +
  geom_abline(slope=1, size=0.2358491) +
  geom_point(size=1, aes(norms_valence, rate_valence), colour = defgrey) + 
  geom_point(data = c_rate_v, colour = "black", aes(x=norms_valence_mean, y=rate_valence_mean), size = 1.5) +
  geom_errorbarh(data = c_rate_v, colour = "black", aes(x = norms_valence_mean, xmin = norms_valence_mean - norms_valence_se, xmax = norms_valence_mean + norms_valence_se, y = rate_valence_mean),  size=0.2358491, height = 0.1) +
  geom_errorbar(data = c_rate_v, colour = "black", aes(ymin = rate_valence_mean - rate_valence_se, ymax = rate_valence_mean + rate_valence_se, x = norms_valence_mean),  size=0.2358491, width=0.1) +
  facet_wrap(~valence)+
  #geom_smooth(method = "lm", se = T, colour = "black", size=0.2358491) + 
  mytheme + #mytheme #theme_classic() +
  scale_x_continuous(name = "normative rating - valence",
                           breaks = seq(1, 8, 2),
                           limits=c(1, 8)) +
  scale_y_continuous(name = "mean individual rating - valence",
                     breaks = seq(1, 8, 1),
                           limits=c(1, 8)) 
cor_val

# ggsave("fig_corval.pdf", plot = cor_val, device = "pdf", path = path_figures,
#   dpi = 300, width = 15, height = 6, units = "cm", limitsize = T)

#----------------------------------------------------------------------------
# 2. Arousal ratings
cor_aro <- ggplot(rate_va) +
  geom_abline(slope=1, size=0.2358491) +
  geom_point(size=1, aes(norms_arousal, rate_arousal), colour = defgrey) + #shape=15
  geom_point(data = c_rate_a, colour = "black", aes(x=norms_arousal_mean, y=rate_arousal_mean), size = 1.5) +
  geom_errorbarh(data = c_rate_a, colour = "black", aes(x = norms_arousal_mean, xmin = norms_arousal_mean - norms_arousal_se, xmax = norms_arousal_mean + norms_arousal_se, y = rate_arousal_mean),  size=0.2358491, height = 0.1) +
  geom_errorbar(data = c_rate_a, colour = "black", aes(ymin = rate_arousal_mean - rate_arousal_se, ymax = rate_arousal_mean + rate_arousal_se, x = norms_arousal_mean),  size=0.2358491, width=0.1) +
  facet_wrap(~valence)+
  #geom_smooth(method = "lm", se = T, colour = "black", size=0.2358491) + 
  mytheme + #mytheme #theme_classic() +
  scale_x_continuous(name = "normative rating - arousal",
                           breaks = seq(2, 7, 1),
                           limits=c(2, 7)) +
  scale_y_continuous(name = "mean individual rating - arousal",
                     breaks = seq(2, 7, 1),
                           limits=c(2, 7)) 
cor_aro

# ggsave("fig_coraro.pdf", plot = cor_aro, device = "pdf", path = path_figures,
#   dpi = 300, width = 15, height = 6, units = "cm", limitsize = T)
#----------------------------------------------------------------------------
```



